### 2018-07-03 firstPaper/figures/gridQuiver.tex
first paper

\begin{tikzpicture}[scale=1.4]
    \node[inner sep = 1pt] (a) at (1,3) {$\bullet$};
    \node[inner sep = 1pt] (b) at (2,3) {$\bullet$};
    \node[inner sep = 1pt] (c) at (1,2) {$\bullet$};
    \node[inner sep = 1pt] (d) at (2,2) {$\bullet$};
    \node[inner sep = 1pt] (e) at (3,2) {$\bullet$};
    \node[inner sep = 1pt] (f) at (4,2) {$\bullet$};
    \node[inner sep = 1pt] (g) at (2,1) {$\bullet$};
    \node[inner sep = 1pt] (h) at (3,1) {$\bullet$};
    \node (x1) at (.2,3) {};
    \node (x2) at (.2,2) {};
    \node (x3) at (1,1.2) {};
    \node (x4) at (1.2,1) {};
    \node (x5) at (2,.2) {};
    \node (x6) at (3,.2) {};
    \node (x7) at (3.8,1) {};
    \node (x8) at (4,1.2) {};
    \node (x9) at (4.8,2) {};
    \node (x10) at (4,2.8) {};
    \node (x11) at (3,2.8) {};
    \node (x12) at (2.8,3) {};
    \node (x13) at (2,3.8) {};
    \node (x14) at (1,3.8) {};
    \path [->, ultra thick] (a) edge (b);
    \path [->, ultra thick] (c) edge (a);
    \path [->, ultra thick] (d) edge (b);
    \path [->, ultra thick] (c) edge (d);
    \path [->, ultra thick] (d) edge (e);
    \path [->, ultra thick] (e) edge (f);
    \path [->, ultra thick] (g) edge (d);
    \path [->, ultra thick] (h) edge (e);
    \path [->, ultra thick] (g) edge (h);
    \path [->] (x1) edge (a);
    \path [->] (x2) edge (c);
    \path [->] (x3) edge (c);
    \path [->] (x4) edge (g);
    \path [->] (x5) edge (g);
    \path [->] (x6) edge (h);
    \path [->] (h) edge (x7);
    \path [->] (x8) edge (f);
    \path [->] (f) edge (x9);
    \path [->] (f) edge (x10);
    \path [->] (e) edge (x11);
    \path [->] (b) edge (x12);
    \path [->] (b) edge (x13);
    \path [->] (a) edge (x14);
    \draw pic[draw, angle eccentricity=1.2, angle radius=.4cm, ultra thick]{angle=c--a--b};
    \draw pic[draw, angle eccentricity=1.2, angle radius=.4cm, ultra thick]{angle=b--d--c};
    \draw pic[draw, angle eccentricity=1.2, angle radius=.4cm, ultra thick]{angle=g--d--e};
    \draw pic[draw, angle eccentricity=1.2, angle radius=.4cm, ultra thick]{angle=e--h--g};
    \draw pic[draw, angle eccentricity=1.2, angle radius=.4cm, ultra thick]{angle=h--e--f};
    \draw pic[draw, angle eccentricity=1.2, angle radius=.4cm]{angle=x14--a--x1};
    \draw pic[draw, angle eccentricity=1.2, angle radius=.4cm]{angle=a--c--x2};
    \draw pic[draw, angle eccentricity=1.2, angle radius=.4cm]{angle=x3--c--d};
    \draw pic[draw, angle eccentricity=1.2, angle radius=.4cm]{angle=d--g--x4};
    \draw pic[draw, angle eccentricity=1.2, angle radius=.4cm]{angle=x5--g--h};
    \draw pic[draw, angle eccentricity=1.2, angle radius=.4cm]{angle=x6--h--x7};
    \draw pic[draw, angle eccentricity=1.2, angle radius=.4cm]{angle=x8--f--x9};
    \draw pic[draw, angle eccentricity=1.2, angle radius=.4cm]{angle=x10--f--e};
    \draw pic[draw, angle eccentricity=1.2, angle radius=.4cm]{angle=x11--e--d};
    \draw pic[draw, angle eccentricity=1.2, angle radius=.4cm]{angle=d--b--x12};
    \draw pic[draw, angle eccentricity=1.2, angle radius=.4cm]{angle=x13--b--a};
\end{tikzpicture}