### 2018-07-04 draft/accordionComplexSurfaces.tex

\documentclass{amsart}

\usepackage[T1]{fontenc}
\usepackage{enumerate, amsmath, amsfonts, amssymb, amsthm, mathrsfs, wasysym, graphics, graphicx, xcolor, url, hyperref, hypcap, xargs, multicol, pdflscape, multirow, hvfloat, array, ae, aecompl, pifont, mathtools, a4wide}
\usepackage{marginnote}
\hypersetup{colorlinks=true, citecolor=darkblue, linkcolor=darkblue}
\usepackage[all]{xy}
%\usepackage[bottom]{footmisc}
\usepackage{tikz}
\usepackage{tkz-graph}
\usetikzlibrary{trees, decorations, decorations.markings, shapes, arrows, matrix, calc, fit, intersections, patterns, angles}
\graphicspath{{figures/}}
\makeatletter\def\input@path{{figures/}}\makeatother
\usepackage{caption}
\captionsetup{width=\textwidth}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% theorems
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem*{theorem*}{Theorem}%[section]

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{question}[theorem]{Question}
\newtheorem{notation}[theorem]{Notation}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{convention}[theorem]{Convention}

% math special letters
\newcommand{\R}{\mathbb{R}} % reals
\newcommand{\N}{\mathbb{N}} % naturals
\newcommand{\Z}{\mathbb{Z}} % integers
\newcommand{\C}{\mathbb{C}} % complex
\newcommand{\I}{\mathbb{I}} % set of integers
\newcommand{\HH}{\mathbb{H}} % hyperplane
\newcommand{\K}{k} % field
\newcommand{\fA}{\mathfrak{A}} % alternating group
\newcommand{\fS}{\mathfrak{S}} % symmetric group
\newcommand{\cA}{\mathcal{A}} % algebra
\newcommand{\cC}{\mathcal{C}} % collection
\newcommand{\cS}{\mathcal{S}} % ground set
\newcommand{\uR}{\underline{R}} % underline set
\newcommand{\uS}{\underline{S}} % underline set
\newcommand{\uT}{\underline{T}} % underline set
\newcommand{\oS}{\overline{S}} % overline set
\newcommand{\ucS}{\underline{\cS}} % underline ground set
\renewcommand{\b}[1]{\mathbf{#1}} % bold letters
\newcommand{\h}{\widehat} % hat letters

% math commands
\newcommand{\set}[2]{\left\{ #1 \;\middle|\; #2 \right\}} % set notation
\newcommand{\bigset}[2]{\big\{ #1 \;\big|\; #2 \big\}} % big set notation
\newcommand{\Bigset}[2]{\Big\{ #1 \;\Big|\; #2 \Big\}} % Big set notation
\newcommand{\setangle}[2]{\left\langle #1 \;\middle|\; #2 \right\rangle} % set notation
\newcommand{\ssm}{\smallsetminus} % small set minus
\newcommand{\dotprod}[2]{\left\langle \, #1 \; \middle| \; #2 \, \right\rangle} % dot product
\newcommand{\symdif}{\,\triangle\,} % symmetric difference
\newcommand{\one}{{1\!\!1}} % the all one vector
\newcommand{\eqdef}{\mbox{\,\raisebox{0.2ex}{\scriptsize\ensuremath{\mathrm:}}\ensuremath{=}\,}} % :=
\newcommand{\defeq}{\mbox{~\ensuremath{=}\raisebox{0.2ex}{\scriptsize\ensuremath{\mathrm:}} }} % =:
\newcommand{\simplex}{\triangle} % simplex
\renewcommand{\implies}{\Rightarrow} % imply sign
\newcommand{\transpose}[1]{{#1}^t} % transpose matrix

% operators
\DeclareMathOperator{\conv}{conv} % convex hull
\DeclareMathOperator{\vect}{vect} % linear span
\DeclareMathOperator{\cone}{cone} % cone hull
\DeclareMathOperator{\inv}{inv} % inversions
\DeclareMathOperator{\ascents}{asc} % ascents
\DeclareMathOperator{\descents}{des} % descents

% others
\newcommand{\fref}[1]{Figure~\ref{#1}} % reference figures
\newcommand{\ie}{\textit{i.e.}~} % id est
\newcommand{\eg}{\textit{e.g.}~} % exempli gratia
\newcommand{\Eg}{\textit{E.g.}~} % exempli gratia
\newcommand{\apriori}{\textit{a priori}} % a priori
\newcommand{\viceversa}{\textit{vice versa}} % vice versa
\newcommand{\versus}{\textit{vs.}~} % versus
\newcommand{\aka}{\textit{a.k.a.}~} % also known as
\newcommand{\perse}{\textit{per se}} % per se
\newcommand{\ordinal}{\textsuperscript{th}} % th for ordinals
\newcommand{\ordinalst}{\textsuperscript{st}} % st for ordinals
\definecolor{darkblue}{rgb}{0,0,0.7} % darkblue color
\definecolor{green}{RGB}{57,181,74} % green color
\definecolor{violet}{RGB}{147,39,143} % violet color
\newcommand{\red}{\color{red}} % red command
\newcommand{\green}{\color{green}} % green command
\newcommand{\darkblue}{\color{darkblue}} % darkblue command
\newcommand{\defn}[1]{\textsl{\darkblue #1}} % emphasis of a definition
\newcommand{\para}[1]{\medskip\noindent\textbf{#1.}} % paragraph
\renewcommand{\topfraction}{1} % possibility to have one page of pictures
\renewcommand{\bottomfraction}{1} % possibility to have one page of pictures
%\renewcommand\labelitemi{$\diamond$} % redefine itemize default symbol
\newcommand{\ex}{_{\textrm{exm}}} % examples
\newcommand{\pa}{_{\textrm{pa}}} % path

% marginal comments
\usepackage{todonotes}
\newcommand{\yann}[1]{\textbf{#1 --- Y.}} % {\todo[color=red!30]{#1 \\ \hfill --- Y.}}
\newcommand{\vincent}[1]{\todo[color=blue!30]{#1 \\ \hfill --- V.}}
\newcommand{\pierreguy}[1]{\todo[color=green!30]{#1 \\ \hfill --- PG.}}

% SPECIFIC NON-KISSING COMPLEX

% COMBINATORICS
% quiver
\newcommand{\blossom}{^\text{\ding{96}}} % blossom
\newcommand{\blinkers}[1]{_{\LEFTCIRCLE \!\! #1 \!\! \RIGHTCIRCLE}} % blossom
\newcommand{\indeg}{\mathrm{indeg}} % in-degree
\newcommand{\outdeg}{\mathrm{outdeg}} % out-degree
%\newcommand{\tai}{\mathsf{tai}} % target incompatible
%\newcommand{\tac}{\mathsf{tac}} % target compatible
%\newcommand{\soi}{\mathsf{soi}} % source incompatible
%\newcommand{\soc}{\mathsf{soc}} % source compatible
%\newcommand{\cso}{\mathsf{cso}} % cosource
%\newcommand{\cta}{\mathsf{cta}} % cotarget
\newcommand{\Enrs}[1]{E_{nr}^{s}(#1)}
\newcommand{\Ers}[1]{E_{r}^{s}(#1)}
\newcommand{\Enrt}[1]{E_{nr}^{t}(#1)}
\newcommand{\Ert}[1]{E_{r}^{t}(#1)}

% walks
\newcommand{\strings}{\mathcal{S}} % strings
\newcommand{\distinguishableStrings}{\mathcal{S}_\mathrm{dist}} % distinguishable strings
\newcommand{\lfStrings}{\underline{\mathcal{S}}} % loop-free strings
\newcommand{\walks}{\mathcal{W}} % walks
\newcommand{\straightWalks}{\mathcal{W}_\mathrm{str}} % straight walks
\newcommand{\bendingWalks}{\mathcal{W}_\mathrm{bend}} % bending walks
\newcommand{\NKWalks}{\mathcal{W}_\mathrm{nk}} % non-kissing walks
\newcommand{\peaks}[1]{\mathsf{peaks}(#1)} % peaks
\newcommand{\deeps}[1]{\mathsf{deeps}(#1)} % deeps
\newcommand{\distinguishedWalk}[2]{\mathsf{dw}(#1,#2)} % distinguished walk
\newcommand{\distinguishedArrows}[2]{\mathsf{da}(#1,#2)} % distinguished arrows
\newcommand{\distinguishedString}[2]{\mathsf{ds}(#1,#2)} % distinguished string
\newcommand{\distinguishedSign}[2]{\varepsilon(#1,#2)} % distinguished sign

% non-kissing complex, lattice, flip graph
\newcommand{\kn}{\kappa} % kissing number
\newcommand{\KN}{\textsc{kn}} % Kissing Number
\newcommandx{\NKC}[1][1=\bar Q]{\mathcal{K}_{\mathrm{nk}}(#1)} % non-kissing complex
\newcommandx{\RNKC}[1][1=\bar Q]{\mathcal{C}_{\mathrm{nk}}(#1)} % non-kissing complex
\newcommandx{\NKL}[1][1=\bar Q]{\mathcal{L}_{\mathrm{nk}}(#1)} % non-kissing lattice
\newcommandx{\NKG}[1][1=\bar Q]{\mathcal{G}_{\mathrm{nk}}(#1)} % non-kissing flip graph
\newcommandx{\NFC}[1][1=\bar Q]{\mathcal{C}_{\mathrm{nf}}(#1)} % non-friendly complex
\newcommand{\peak}{\mathrm{peak}} % bottom dissection
\newcommand{\deep}{\mathrm{deep}} % top dissection
\newcommand{\reversed}[1]{#1^{\mathrm{rev}}} % reverse all arrows
\renewcommand{\top}{\mathrm{top}} % top
\newcommand{\bottom}{\mathrm{bot}} % bottom
\newcommand{\walk}{\operatorname{\omega}} % walk (of a curve)

% accordion and slalom complexes
\newcommand{\surface}{\mathcal{S}} % an orientable surface
\newcommand{\dual}{^*} % duality
\newcommand{\dissection}{\mathrm{D}} % dissection of a marked surface
\newcommand{\vertices}{\mathcal{V}} % vertices of a dissection
\newcommand{\edges}{\mathcal{E}} % edges of a dissection
\newcommand{\faces}{\mathcal{F}} % faces of a dissection
\renewcommandx{\AC}[1][1=\dissection]{\mathcal{K}_{\mathrm{acc}}(#1)} % accordion complex
\newcommandx{\RAC}[1][1=\dissection]{\mathcal{C}_{\mathrm{acc}}(#1)} % reduced accordion complex
\newcommandx{\SC}[1][1=\dissection\dual]{\mathcal{K}_{\mathrm{sla}}(#1)} % slalom complex
\newcommandx{\RSC}[1][1=\dissection\dual]{\mathcal{C}_{\mathrm{sla}}(#1)} % reduced slalom complex
\newcommandx{\NCC}[1][1={\dissection, \dissection\dual}]{\mathcal{K}_{\mathrm{nc}}(#1)} % non-crossing complex
\newcommandx{\RNCC}[1][1={\dissection, \dissection\dual}]{\mathcal{C}_{\mathrm{nc}}(#1)} % non-crossing complex
\newcommand{\curveof}{\operatorname{\gamma}} % curve (of a walk)

% lattices
\newcommand{\meet}{\wedge} % meet
\newcommand{\join}{\vee} % join
\newcommand{\bigMeet}{\bigwedge} % meet
\newcommand{\bigJoin}{\bigvee} % join
\newcommand{\closure}[1]{#1^{\mathrm{cl}}} % closure operator
\newcommand{\coclosure}[1]{#1^{\mathrm{cocl}}} % coclosure operator
\newcommand{\uclosure}[1]{#1^{\underline{\mathrm{cl}}}} % loop free closure operator
\newcommand{\Bicl}[1]{\mathsf{Bic}(#1)} % biclosed sets
\newcommand{\projDown}{\pi_\downarrow} % down projection map
\newcommand{\projUp}{\pi^\uparrow} % up projection map
\newcommand{\JI}{\mathsf{JI}} % join-irreducible
\newcommand{\MI}{\mathsf{MI}} % meet-irreducible
\newcommand{\Cong}{\mathsf{Cong}} % congruence lattice
\newcommand{\con}{\mathrm{con}} % congruence contracting a cover relation
\newcommand{\ji}{\mathsf{ji}} % join irreducible
\newcommand{\mi}{\mathsf{mi}} % meet irreducible

% GEOMETRY
% g-vectors
\newcommand{\gvector}[1]{\mathbf{g}(#1)} % g-vector of the path #1
\newcommand{\gvectors}[1]{\mathbf{g}(#1)} % g-vectors of the paths #1
\newcommandx{\gvectorFan}[1][1=\bar Q]{\mathcal{F}^\mathbf{g}(#1)} % g-vector fan
% c-vectors
\newcommand{\cvector}[2]{\mathbf{c}(#1 \in #2)} % c-vector of the path #1 in the facet #2
\newcommand{\cvectors}[1]{\mathbf{c}(#1)} % c-vectors of the paths #1
\newcommandx{\allcvectors}[1][1=\bar Q]{\mathbf{C}(#1)} % all c-vectors with respect to the initial cluster #1
\newcommandx{\cvectorFan}[1][1=\bar Q]{\mathcal{F}^\mathbf{c}(#1)} % fan of hyperplanes orthogonal to all c-vectors
% points, hyperplanes, half-spaces
\newcommand{\point}[1]{\mathbf{p}(#1)} % vertex of the grid associahedron corresponding to the cluster #1
\newcommand{\ray}{\mathbf{r}} % ray
\newcommand{\rays}{\mathbf{R}} % rays
\newcommand{\hs}{\mathbf{H}^{\le}} % half space
\newcommand{\HS}[1]{\mathbf{H}^{\le}(#1)} % half space
\newcommand{\hyp}{\mathbf{H}^{=}} % hyperplane
\newcommand{\Hyp}[1]{\mathbf{H}^{=}(#1)} % hyperplane
\newcommand{\fix}[1]{\mathrm{Fix}(#1)} % fix space
% polytopes
\newcommandx{\Asso}[2][1=\bar Q,2={}]{\mathsf{Asso}^{#2}(#1)} % associahedron
\newcommandx{\Zono}[2][1=\bar Q,2={}]{\mathsf{Zono}^{#2}(#1)} % zonotope
\newcommand{\Fan}{\mathcal{F}} % fan
\newcommand{\multiplicityVector}{\b{m}} % multiplicity vector

% ALGEBRA
\newcommand{\Hom}[1]{\operatorname{{\rm Hom}}_{#1}}
\newcommand{\Ext}[1]{\operatorname{{\rm Ext}}_{#1}}
\newcommandx{\AR}[1][1=\bar Q]{\mathrm{AR}(#1)} % Auslander-Reiten quiver
\newcommandx{\tTC}[1][1=\bar Q]{\mathcal{K}^{\textrm{s$\tau$-tilt}}(#1)} % tau-tilting complex
\newcommand{\rep}{\operatorname{{\rm rep}}}
\newcommand{\proj}{\operatorname{{\rm proj}}}
\newcommand{\koszul}{^!} % duality

%% formating the part command
%\makeatletter
%\def\part{\@startsection{part}{1}%
%\z@{.7\linespacing\@plus\linespacing}{.8\linespacing}%
%{\LARGE\sffamily\centering}}
%\@addtoreset{section}{part}
%\makeatother
%\renewcommand{\thesection}{\arabic{part}.\arabic{section}}

% formating the table of contents
\makeatletter
\def\l@section{\@tocline{1}{2pt}{0pc}{}{}}
\makeatother
\let\oldtocpart=\tocpart
\renewcommand{\tocpart}[2]{\bf\large\oldtocpart{#1}{#2}}
\let\oldtocsection=\tocsection
\renewcommand{\tocsection}[2]{\bf\oldtocsection{#1}{#2}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title[Non-kissing and non-crossing complexes for locally gentle algebras]{Non-kissing and non-crossing complexes \\ for locally gentle algebras}

\thanks{YP, VP and PGP were partially supported by the French ANR grant SC3A~(15\,CE40\,0004\,01).}

\author{Yann Palu}
\address[Yann Palu]{LAMFA, Universit\'e Picardie Jules Verne, Amiens}
\email{yann.palu@u-picardie.fr}
\urladdr{\url{http://www.lamfa.u-picardie.fr/palu/}}

\author{Vincent Pilaud}
\address[Vincent Pilaud]{CNRS \& LIX, \'Ecole Polytechnique, Palaiseau}
\email{vincent.pilaud@lix.polytechnique.fr}
\urladdr{\url{http://www.lix.polytechnique.fr/~pilaud/}}

\author{Pierre-Guy Plamondon}
\address[Pierre-Guy Plamondon]{Laboratoire de Math\'ematiques d'Orsay, Universit\'e Paris-Sud, CNRS, Universit\'e Paris-Saclay}
\email{pierre-guy.plamondon@math.u-psud.fr}
\urladdr{\url{https://www.math.u-psud.fr/~plamondon/}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{abstract}
We consider three combinatorial models for the support $\tau$-tilting complex of a locally gentle bound quiver: the accordion complex, the slalom complex, and the non-kissing complex. We unify these models in the same picture.
\end{abstract}

%\vspace*{.3cm}

\maketitle

Possible titles, depending on what we want to insist on:
\begin{enumerate}
\item Accordion / slalom / non-kissing / support $\tau$-tilting complexes all coincide for gentle algebras
\item Non-kissing and non-crossing complexes for gentle algebras
\item Combinatorial models for support $\tau$-tilting on gentle algebras: accordion, slalom and non-kissing complexes
\end{enumerate}

\section{Non-kissing complex}

\subsection{Locally gentle bound quivers and their blossoming quivers}

%We consider a \defn{bound quiver}~$\bar Q = (Q,I)$, formed by a quiver~$Q$ and an admissible ideal~$I$ of the path algebra~$kQ$, so that the quotient algebra~$kQ/I$ is finite dimensional.
We consider a \defn{bound quiver}~$\bar Q = (Q,I)$, formed by a quiver~$Q$ and an ideal~$I$ of the path algebra~$kQ$ such that $I$ is generated by linear combinations of paths of length at least two.
Note that we \textbf{do not require} that the quotient algebra~$kQ/I$ be finite dimensional.

\begin{definition}[\cite{ButlerRingel}]
\label{def:gentleQuiver}
A \defn{locally gentle bound quiver}~$\bar Q \eqdef (Q,I)$ is a bound quiver such that
\begin{itemize}
\item each vertex~$v \in Q_0$ has at most two incoming and two outgoing arrows,
\item the ideal~$I$ is generated by paths of length exactly two,
\item for any arrow~$\beta \in Q_1$, there is at most one arrow~$\alpha \in Q_1$ such that~$t(\alpha) = s(\beta)$ and~${\alpha\beta\notin I}$ (resp.~$\alpha\beta \in I$) and at most one arrow~$\gamma \in Q_1$ such that~$t(\beta) = s(\gamma)$~and~${\beta\gamma\notin I}$~(resp.~${\beta\gamma \in I}$).
\end{itemize}
The algebra~$kQ/I$ is called a \defn{locally gentle algebra}.
A \defn{gentle bound quiver} is a locally bound quiver~$\bar Q$ such that the algebra $kQ/I$ is finite-dimensional; the algebra is then called a \defn{gentle algebra}.
\end{definition}

\begin{definition}[\cite{PaluPilaudPlamondon}]
\label{def:blossomingQuiver}
A locally gentle bound quiver~$\bar Q$ is \defn{complete} if any vertex~$v \in Q_0$ is incident to either one ($v$ is a leaf) or four arrows ($v$ is an internal vertex).
The \defn{pruned subquiver} of a quiver $\bar Q$ is the locally gentle quiver obtained by deleting all leaves of~$\bar Q$ (degree one vertices) and their incident arrows.
The \defn{blossoming quiver} of a locally gentle bound quiver~$\bar Q$ is the complete locally gentle bound quiver~$\bar Q\blossom$ whose pruned subquiver is~$\bar Q$.
The vertices of~$Q\blossom_0 \ssm Q_0$ are called \defn{blossom vertices}, and the arrows~$Q\blossom_1 \ssm Q_1$ are called \defn{blossom arrows}.
\end{definition}

Examples are represented on \fref{fig:quivers}.

\begin{figure}[t]
	\capstart
	\centerline{\includegraphics[scale=.6]{quivers}}
	\caption{Some locally gentle quivers and their blossoming quivers. Initial vertices are solid and labeled while blossom vertices are hollow, and initial arcs are bold while blossom arcs are thin.}
	\label{fig:quivers}
\end{figure}

\vincent{Should we introduce Koszul duality here and say what it does on quivers?}

\subsection{Strings and walks}

\begin{definition}
\label{def:string}
Let~$\bar Q = (Q,I)$ be a locally gentle bound quiver.
A \defn{string} for~$\bar Q$ is a word of the form
\(
\rho = \alpha_1^{\varepsilon_1}\alpha_2^{\varepsilon_2}\cdots \alpha_\ell^{\varepsilon_\ell},
\)
where
	\begin{itemize}
	\item $\alpha_i \in Q_1$ and~$\varepsilon_i \in \{-1,1\}$ for all~$i \in [\ell]$,
	\item $t(\alpha_i^{\varepsilon_i}) = s(\alpha_{i+1}^{\varepsilon_{i+1}})$ for all~$i \in [\ell-1]$,
	\item there is no path~$\pi \in I$ such that~$\pi$ or~$\pi^{-1}$ appears as a factor of~$\rho$, and
	\item $\rho$ is reduced, in the sense that no factor~$\alpha\alpha^{-1}$ or~$\alpha^{-1}\alpha$ appears in~$\rho$, for~$\alpha \in Q_1$.
	\end{itemize}
The integer~$\ell$ is called the \defn{length} of the string~$\rho$.
We let~$s(\rho) \eqdef s(\alpha_1^{\varepsilon_1})$ and~$t(\rho) \eqdef t(\alpha_\ell^{\varepsilon_\ell})$ denote the source and target of~$\rho$.
For each vertex~$v \in Q_0$, there is also a \defn{string of length zero}, denoted by~$\varepsilon_v$, that starts and ends at~$v$.
%We denote by~$\strings(\bar Q)$ the set of strings on~$\bar Q$.
%We usually denote strings by~$\rho, \sigma, \tau$.
We often implicitly identify the two inverse strings~$\rho$ and~$\rho^{-1}$, and we let~$\strings^\pm(\bar Q) \eqdef \set{\{\rho, \rho^{-1}\}}{\rho \in \strings(\bar Q)}$ denote the collection of all \defn{undirected strings} of~$\bar Q$.

%We call \defn{negative simple string} a formal word of length zero of the form $-v$, where $v$ is any vertex of $Q_0$. We denote by
%\(
%\strings^\pm_{\ge -1}(\bar Q) \eqdef \strings^\pm(\bar Q) \sqcup \set{-v}{v \in Q_0}
%\)
%the set of \defn{almost positive unoriented strings}.
%Note that~$\varepsilon_v$ and~$-v$ are different almost positive strings.
%\end{enumerate}
\end{definition}

\begin{definition}
\label{def:walk}
A \defn{walk} of a locally gentle quiver~$\bar Q$ is a maximal string of its blossoming quiver~$\bar Q \blossom$, meaning that at each end it either reaches a blossom vertex of~$\bar Q \blossom$ or enters an infinite oriented cycle.
We denote by~$\walks(\bar Q)$ the set of all walks on~$\bar Q$.
As for strings, we implicitly identify the two inverse walks~$\omega$ and~$\omega^{-1}$, and we let~$\walks^\pm(\bar Q) \eqdef \set{\{\omega, \omega^{-1}\}}{\omega \in \walks(\bar Q)}$ denote the set of \defn{undirected walks} of~$\bar Q$.
\end{definition}

\begin{definition}
\label{def:substrings}
A \defn{substring} of a walk~$\omega = \alpha_1^{\varepsilon_1} \dots \alpha_\ell^{\varepsilon_\ell} \in \walks(\bar Q)$ is a string~$\sigma = \alpha_i^{\varepsilon_i} \dots \alpha_j^{\varepsilon_j}$ of~$\strings(\bar Q)$ for some~$1 < i \le j < \ell$.
Note that 
\begin{itemize}
\item the endpoints of~$\sigma$ are not allowed to be the blossom endpoints of~$\omega$,
\item the position of~$\sigma$ as a factor of~$\omega$ matters (the same string at a different position is considered a different substring),
\item the string~$\varepsilon_v$ is a substring of~$\omega$ for each occurence of~$v$ as a vertex of~$\omega$.
\end{itemize}
The substring~$\sigma = \alpha_i^{\varepsilon_i} \dots \alpha_j^{\varepsilon_j}$ is \defn{at the bottom} (resp.~\defn{on top}) of the walk~$\omega = \alpha_1^{\varepsilon_1} \dots \alpha_\ell^{\varepsilon_\ell}$ if~$\varepsilon_{i-1} = 1$ and~$\varepsilon_{j+1} = -1$ (resp.~if~$\varepsilon_{i-1} = -1$ and~$\varepsilon_{j+1} = 1$), that is, the two arrows of~$\omega$ incident to the endpoints of~$\sigma$ point towards~$\sigma$ (resp.~outwards from~$\sigma$).
We let~$\Sigma(\omega)$ be the set of substrings of~$\omega$ and~$\Sigma_\bottom(\omega)$ and~$\Sigma_\top(\omega)$ be the sets of bottom and top substrings of~$\omega$ respectively.
We use the same notations for undirected walks (of course, substrings of an undirected walk are undirected).
\end{definition}

\begin{definition}
\label{def:straightBended}
A \defn{peak} (resp.~\defn{deep}) of a walk~$\omega$ is a substring of~$\omega$ of length zero which is on top (resp.~at the bottom of~$\omega$).
A \defn{corner} is either a peak or a deep (in other words, it is a vertex of~$\omega$ where the arrows change direction).
A walk~$\omega$ is \defn{straight} if it has no corner (\ie if~$\omega$ or~$\omega^{-1}$ is a path in~$\bar Q\blossom$), and \defn{bending} otherwise. Let $\straightWalks(\bar Q)$ and~$\bendingWalks(\bar Q)$ denote the sets of straight and bending walks of~$\bar Q$ and let~$\straightWalks^\pm(\bar Q) \eqdef \set{\{\omega, \omega^{-1}\}}{\omega \in \straightWalks(\bar Q)}$ and~$\bendingWalks^\pm(\bar Q) \eqdef \set{\{\omega, \omega^{-1}\}}{\omega \in \bendingWalks(\bar Q)}$.
A \defn{peak walk} (resp.~\defn{deep walk}) is a walk with a unique corner (in other words, it switches orientation only once), which is a peak (resp.~deep). For~$v \in Q_0$, we denote by~$v_\peak$ the peak walk with peak at~$v$ and by~$v_\deep$ the deep walk with deep~at~$v$.
\end{definition}

\subsection{Non-kissing complex}

\begin{definition}
\label{def:kiss}
Let~$\omega,\omega' \in \walks^\pm(\bar Q)$ be two undirected walks on~$\bar Q$.
We say that~$\omega$ \defn{kisses}~$\omega'$ if~$\Sigma_\top(\omega) \cap \Sigma_\bottom(\omega') \ne \varnothing$.
See \fref{fig:kissingCrossing}.
% \ie if there exist a common substring~$\sigma \in \strings^\pm(\bar Q)$ of~$\omega$ and~$\omega'$ such that the arrows of~$\omega$ incident to~$\sigma$ are both outgoing while the arrows of~$\omega'$ incident to~$\sigma$ are both incoming.
We say that~$\omega$ and~$\omega'$ are \defn{kissing} if~$\omega$ kisses~$\omega'$ or~$\omega'$ kisses~$\omega$ (or both).
Note that we authorize the situation where~$\sigma$ is reduced to a vertex~$v$, meaning that~$v$ is a peak of~$\omega$ and a deep of~$\omega'$.
Observe also that~$\omega$ can kiss~$\omega'$ several times, that~$\omega$ and~$\omega'$ can mutually kiss, and that~$\omega$ can kiss itself.
We denote by~$\NKWalks(\bar Q)$ the walks on~$\bar Q$ which are not self-kissing and by~$\NKWalks^\pm(\bar Q) \eqdef \set{\{\omega, \omega^{-1}\}}{\omega \in \NKWalks(\bar Q)}$ their undirected version.
\end{definition}

\begin{figure}
	\capstart
	\centerline{\includegraphics[scale=1]{kissingCrossing}}
	\caption{A schematic representation of two kissing walks (left) and two crossing walks~(right): $\omega$~kisses~$\omega'$ (left) while~$\mu$~crosses~$\nu$ (right). For the crossing walks, we have~$\mu \prec_\gamma \nu$ while~$\nu \prec_\delta \mu$.}
	\label{fig:kissingCrossing}
\end{figure}

\begin{definition}
\label{def:nKc}
The \defn{non-kissing complex} of~$\bar Q$ is the simplicial complex~$\NKC$ whose faces are the collections of pairwise non-kissing walks of~$\NKWalks^\pm(\bar Q)$.
Note that self-kissing walks never appear in~$\NKC$ by definition.
In contrast, no straight walk can kiss another walk by definition, so that they appear in all facets of~$\NKC$.
We thus consider the \defn{reduced non-kissing complex}~$\RNKC$ to be the deletion of all straight walks from~$\NKC$.
\end{definition}

\begin{theorem}[{\cite[Cor.~2.27 \& Cor.~2.33]{PaluPilaudPlamondon}}\footnote{As the main objective of~\cite[Cor.~2.27 \& Cor.~2.33]{PaluPilaudPlamondon} was a combinatorial model for support $\tau$-tilting complexes of gentle algebras, we did require that the ideal~$I$ was admissible so that~$kQ/I$ was finite dimensional. However, the same proof directly applies in the generality of this paper.}]
The reduced non-kissing complex~$\RNKC$ is pure of dimension~$|Q_0|$ (\ie all maximal faces have dimension~$|Q_0|$) and thin (\ie all codimension~$1$ faces are contained in exactly two facets).
\end{theorem}

This result follows from the existence of flips in non-kissing facets.
This operation is illustrated in \fref{fig:flip}.

\begin{figure}[t]
	\capstart
	\centerline{\includegraphics[scale=1]{flip}}
	\caption{A schematic representation of flips in the non-kissing complex. The flip exchanges the walk~$\omega = \rho \sigma \tau$ in the facet~$F$ (left) with the walk~$\omega' = \rho' \sigma \tau'$ in the facet~$F'$ (middle), using the walks~$\mu = \rho' \sigma \tau$ and~$\nu = \rho \sigma \tau'$. The walk~$\omega'$ kisses~$\omega$ but no other walk of~$F$ (right).}
	\label{fig:flip}
\end{figure}

\section{Accordion complex, slalom complex, and non-crossing complex}

\subsection{Dual dissections of a surface}

\begin{definition}
A \defn{marked surface}~$\bar\surface \eqdef (\surface, M)$ is an orientable surface~$\surface$ with boundaries, together with a set~$M$ of marked points which can be on the boundary of~$\surface$ or not.
For~$V \subset \surface$,
\begin{itemize}
\item a \defn{$V$-arc} on~$\bar\surface$ is a curve on~$\surface$ connecting two points of~$V$ and whose interior is disjoint from~$M$ and the boundary of~$\surface$.
\item a \defn{$V$-curve} on~$\bar\surface$ is a curve on~$\surface$ which at each end either reaches a point of~$V$ or infinitely circles around and finally reaches a puncture of~$M$.
\end{itemize}
As usual, curves and arcs are considered up to homotopy relative to their endpoints in~$\surface \ssm M$, and curves homotopic to a boundary are not allowed.
\end{definition}

\begin{definition}
Two curves or arcs \defn{cross} when they intersect in~their~interior.
We will always assume that collections of arcs on a surface are in minimal position, in the sense that they cross each other transversaly, and the number of crossings is minimal.
This is always possible up to homotopy. \pierreguy{Reference?}
\end{definition}

\begin{definition}
A \defn{dissection} of~$\bar\surface$ is a collection~$\dissection$ of pairwise non-crossing arcs on~$\bar\surface$.
The \defn{edges} of~$\dissection$ are its arcs together with the boundary arcs of~$\bar\surface$.
The \defn{faces} of~$\dissection$ are the connected components of the complement of the union of the edges of~$\dissection$ in the surface~$\surface$.
We denote by~$\vertices(\dissection)$, $\edges(\dissection)$ and~$\faces(\dissection)$ the sets of vertices, edges and faces of~$\dissection$ respectively.
The dissection~$\dissection$ is \defn{cellular} if all its faces are disks.
For~$V \subseteq M$, a \defn{$V$-dissection} is a dissection with only $V$-arcs.
\end{definition}

\begin{convention}
\textbf{All throughout the paper, all dissections are considered cellular.}
The case of non-cellular dissections will be discussed in Section~\ref{sec:nonCellularDissections}.
\end{convention}

\begin{definition}
Consider a marked surface~$\bar\surface = (\surface, V \sqcup V\dual)$, where~$V$ and~$V\dual$ are two disjoint sets of marked points so that the points of~$V$ and~$V\dual$ that are on the boundary of~$\surface$ alternate.
A cellular $V$-dissection~$\dissection$ of~$\bar\surface$ and a cellular $V\dual$-dissection~$\dissection\dual$ of~$\bar\surface$ are \defn{dual cellular dissections} if there are pairs of mutually inverse bijections~$V\dual \leftrightarrow \faces(\dissection)$ and~$V \leftrightarrow \faces(\dissection\dual)$, denoted~$\dual$ in both directions, such that~$\dissection$ has an edge joining its vertices~$u,v \in V$ and separating its faces~$f,g \in  \faces(\dissection)$ if and only if~$\dissection\dual$ has an edge joining its vertices~$f\dual, g\dual \in V\dual$ and separating its faces~$u\dual, v\dual \in \faces(\dissection\dual)$.
\end{definition}

Note that contrarily to the usual conventions, the dual vertex~$f\dual$ of a face~$f$ of~$\dissection$ is not always in the interior of the face~$f$.
More precisely, there are two situations:
\begin{itemize}
\item if a face~$f$ has no edge on the boundary of~$\surface$, its dual vertex~$f\dual$ lies in the interior~of~$f$,
\item if a face~$f$ has an edge on the boundary of~$\surface$, then it has exactly one such edge and its dual vertex~$f\dual$ lies on this edge.
\end{itemize}
In fact, the second point forces the following characterization of the cellular dissections that admit a dual cellular dissection.

\begin{proposition}
\label{prop:conditionsDualDissections}
For a cellular $V$-dissection~$\dissection$ of a marked surface~$\bar\surface = (\surface, V \sqcup V\dual)$, the following assertions are equivalent:
\begin{enumerate}[(i)]
\item there exists a cellular $V\dual$-dissection~$\dissection\dual$ of~$\bar\surface$ such that~$\dissection$ and~$\dissection\dual$ are dual cellular dissections,
\item each face of~$\dissection$ contains exactly one point of~$V\dual$.
\end{enumerate}
Moreover, the cellular dissection~$\dissection\dual$ is uniquely determined.
\end{proposition}

\begin{proof}
We mimick the proof of~\cite[Prop.~1.12]{OpperPlamondonSchroll}.
%Assume first that~$\dissection$ and~$\dissection\dual$ are dual cellular dissections. If a face~$f$ of~$\dissection$ had two boundary arcs~$a, b$, then it would contain two points of~$M\dual$, one on~$a$ and one on~$b$.
%Conversely, if each face of~$\dissection$ has at most one boundary arc, we construct the set of points~$M\dual$ with one point on each boundary arc of~$\surface$ and one point in the interior of each interior face of~$\dissection$. Clearly, $M\dual$ has one point~$f\dual$ in each face~$f$ of~$\dissection$. We then construct half-arcs of~$\dissection\dual$ in each face~$f$ of~$\dissection$ by joining the dual point~$f\dual$ to the middle of each boundary arc of~$f$ that is not a boundary arc of~$\surface$. For each arc~$a$ of~$\dissection$, the half-arcs incident to the middle of~$a$ in the two faces of~$\dissection$ containing~$a$ form the dual arc~$a\dual$ of~$a$. All these dual arcs do not cross (as the half-arcs do not cross in each face of~$\dissection$) and form the dual cellular dissection~$\dissection\dual$ of~$\dissection$.
%More details can be found in~\cite{OpperPlamondonSchroll}.
The direct implication is immediate by definition.
Assume conversely that each face~$f$ of~$\dissection$ contains exactly one point~$f\dual$ of~$V\dual$.
We construct half-edges of~$\dissection\dual$ in each face~$f$ of~$\dissection$ by joining the dual point~$f\dual$ to the middle of each boundary edge of~$f$ that is not a boundary edge of~$\surface$.
For each edge~$a$ of~$\dissection$, the half-edges incident to the middle of~$a$ in the two faces of~$\dissection$ containing~$a$ form the dual edge~$a\dual$ of~$a$.
All these dual edges do not cross (as the half-edges do not cross in each face of~$\dissection$) and form the dual cellular dissection~$\dissection\dual$~of~$\dissection$.
\end{proof}

%\begin{remark}
%In Proposition~\ref{prop:conditionsDualDissections}, note that although the set~$M\dual$ of dual marked points is unique, the dual dissection~$\dissection\dual$ is not when~$\dissection$ is not cellular.
%\end{remark}

Some examples of dual cellular dissections on different surfaces are represented in \fref{fig:dissections}.

\begin{figure}[t]
	\capstart
	\centerline{\includegraphics[scale=.7]{dissections}}
	\caption{Some pairs of dual dissections on different surfaces.}
	\label{fig:dissections}
\end{figure}

\begin{definition}
Interior points, blossom points.
Half-arcs.

We consider a set~$B$ of points on the boundary of the surface~$\surface$ such that~$B$ and~$M \cup M\dual$ alternate along the boundary of~$\surface$.
The points of~$B$ are called the \defn{blossom points}.
\end{definition}

\subsection{Accordion complex, slalom complex, and non-crossing complex}
Let~$\bar\surface = (\surface, V \sqcup V\dual)$ be a surface with two disjoint sets~$V$ and~$V\dual$ of marked points, and let~$B$ be the corresponding blossom points.
%
%\begin{definition}
%A \defn{$B$-curve} is a curve on~$\surface$ which at each end either reaches a point of~$B$ or infinitely circles around and finally reaches a puncture of~$V \sqcup V\dual$.
%Curves are considered up to homotopy relative to their endpoints in~$\surface \ssm (B \sqcup V \sqcup V\dual)$, and curves homotopic to a boundary are not allowed.
%\end{definition}
%
%Let~$I$ denote the interior points and $B$ the blossom points.
We say that a $B$-curve is \defn{external} if it is homotopic to a boundary arc of~$\surface \ssm B$, and \defn{internal} otherwise.
Note that no $B$-curve can cross an external $B$-arc.
Consider two dual cellular dissections~$\dissection$ and~$\dissection\dual$ of~$\bar\surface$.

%With this assumption, each segment of a~$B$-arc that lies in a face~$f$ of the dissection~$\dissection$ has two intersection points (called crossings) with the edges of~$f$.
%Note that the edges of~$f$ may be on the boundary of the surface.

\begin{definition}
\label{def:accordion}
A \defn{$\dissection$-accordion} is a $B$-curve~$\alpha$ of~$\bar\surface$ such that whenever~$\alpha$ meets a face~$f$~of~$\dissection$,
\begin{itemize}
\item it enters crossing an edge~$e_1$ and leaves crossing an edge~$e_2$ (in other words, $\alpha$ is not allowed to circle around~$f\dual$ when~$f\dual$ is a puncture),
\item the two edges~$e_1$ and~$e_2$ of~$f$ crossed by~$\alpha$ are consecutive on the boundary of~$f$,
\item $\alpha$, $e_1$ and~$e_2$ bound a disk inside~$f$ that does not contain~$f\dual{}$.
\end{itemize}
By convention, we also consider that the punctures of~$V\dual$ are $\dissection$-accordions.
\end{definition}

\begin{figure}[t]
	\capstart
	\centerline{\includegraphics[scale=.7]{accordions}}
	\caption{Some accordions.}
	\label{fig:accordions}
\end{figure}

\begin{remark}
In Definition~\ref{def:accordion}, observe that
\begin{itemize}
\item the edges of~$e_1$ and~$e_2$ of~$f$ crossed by~$\alpha$ might coincide, see ...
\item the second condition is automatically satisfied if~$f\dual$ is not a puncture.
\end{itemize}
\end{remark}

\begin{definition}
\label{def:accordionComplex}
The \defn{$\dissection$-accordion complex}~$\AC$ is the simplicial complex of pairwise non-crossing $\dissection$-accordions.
The \defn{reduced $\dissection$-accordion complex}~$\RAC$ is the simplicial complex of pairwise non-crossing internal $\dissection$-accordions.
\end{definition}

\begin{definition}
\label{def:slalom}
A \defn{$\dissection\dual$-slalom} is a $B$-curve~$\alpha$ of~$\bar\surface$ such that, whenever~$\alpha$ crosses an edge~$a$ of~$\dissection$ contained in two faces~$f,g$ of~$\dissection$, the marked points~$f\dual$ and~$g\dual$ lie on opposite sides of~$\alpha$ in the union of~$f$ and~$g$ glued along~$a$.
Here, we consider that~$f\dual$ lies on the right (resp.~left) of~$\alpha$ when~$\alpha$ circles clockwise (resp.~counterclockwise) around~$f\dual$.
By convention, we also consider that the punctures of~$V$ are $\dissection\dual$-slaloms.
\end{definition}

\begin{figure}[t]
	\capstart
	\centerline{\includegraphics[scale=.7]{slaloms}}
	\caption{Some slaloms.}
	\label{fig:slaloms}
\end{figure}

\begin{definition}
\label{def:slalomComplex}
The \defn{$\dissection\dual$-slalom complex}~$\SC$ is the simplicial complex of pairwise non-crossing $\dissection$-slaloms.
The \defn{reduced $\dissection\dual$-slalom complex}~$\RAC$ is the simplicial complex of pairwise non-crossing internal $\dissection\dual$-slaloms.
\end{definition}

%\begin{remark}
%Note that the $\dissection$-accordions do not depend on~$\dissection\dual$, only on~$\dissection$ (and~$M\dual$ which is determined by~$\dissection$ by Proposition~\ref{prop:conditionsDualDissections}).
%Similarly, the $\dissection\dual$-slaloms do not depend on~$\dissection$.
%\end{remark}

\begin{proposition}
\label{prop:accordionsSlaloms}
For two dual cellular dissections~$\dissection$ and~$\dissection\dual$, the $\dissection$-accordions are precisely the $\dissection\dual$-slaloms and the (reduced) $\dissection$-accordion complex coincides with the (reduced) $\dissection\dual$-slalom complex.
\end{proposition}

\begin{proof}
\yann{I have to merge two cases}
Let~$\alpha$ be a~$\dissection$-accordion.
Let~$\beta$,~$\gamma$,~$\delta$ be three consecutive arcs of~$\dissection$ crossed by~$\gamma$.
Let~$m\in\dissection$ be the common endpoint of~$\beta$ and $\gamma$ given by Definition~\ref{def:accordion} and let~$n\in\dissection$ be the common endpoint of~$\gamma$ and~$\delta$.
If $m=n$, then the segment of~$\alpha$ between~$\beta$ and~$\delta$ stays in the same face of~$\dissection\dual$.
Otherwise, when crossing~$\gamma$, the accordion~$\alpha$ leaves the face~$m\dual$ of $\dissection\dual$ and enters the face~$n\dual$.
In the surface obtained by glueing the two cells~$m\dual$ and~$n\dual$ along~$\gamma\dual$, the marked points~$m$ and~$n$ are separated by~$\alpha$.
Conversely, assume that~$\alpha$ is an arc of~$(\surface, B)$ which is not a $\dissection$-accordion.
We are in one of the following cases.

Case 1: The arc~$\alpha$ consecutively crosses two arcs~$\beta$ and~$\gamma$ of~$\dissection$ that do not share a common endpoint.
Let~$f$ be the face of~$\dissection$ that contain the segment of~$\alpha$ between~$\beta$ and~$\gamma$.
Consider the connected component of~$f\setminus\alpha$ that do not contain~$f\dual$.
The boundaty of this component contains an arc~$\delta\in\dissection\setminus\{\beta\}$ that share a common endpoint with~$\beta$.
Let~$m,n$ be the endpoints of~$\delta$.
Then~$\alpha$ cuts the cells $m\dual,n\dual$ glued along~$\delta\dual$ into two connected component, one containing both endpoints of~$\delta$.

Case 2: The arc~$\alpha$ consecutively crosses two arcs~$\beta$ and~$\gamma$ of~$\dissection$ that share a unique common endpoint, but the region delimited by~$\alpha$,~$\beta$ and~$\gamma$ is not a disk.
Since the dissection~$\dissection$ is cellular, that region has to contain some marked point~$m\in\dissection\dual$.
One conclude that~$\alpha$ is not a~$\dissection\dual$-slalom similarly as in the first case, by considering the component of~$m\dual\setminus\alpha$ that do not contain~$m$.

Case 3: The arc~$\alpha$ consecutively crosses two arcs~$\beta$ and~$\gamma$ of~$\dissection$ that share two different endpoints, and none of the two regions delimited by~$\alpha$,~$\beta$ and~$\gamma$ is a disk.
Since the dissection~$\dissection$ is cellular, both regions have to contain a puncture.
This contradicts the assumption that the dissections~$\dissection$ and~$\dissection\dual$ are dual to each other.

Case 4: If~$\alpha$ stays inside a blue cell, then it rolls around a~$\dissection\dual$ puncture.
There is a~$\dissection\dual$-arc~$\beta$ ending in that puncture. When~$\alpha$ crosses~$\beta$, the puncture stays on the same side of~$\alpha$, hence~$\alpha$ is not a~$\dissection\dual$-slalom.
\end{proof}

\begin{definition}
\label{def:noncrossingComplex}
We call \defn{non-crossing complex} of the pair of dual cellular dissections~$(\dissection, \dissection\dual)$ the simplicial complex~$\NCC = \AC = \SC$.
Similarly, the \defn{reduced non-crossing complex} of~$(\dissection, \dissection\dual)$ is~$\RNCC = \RAC = \RSC$.
\end{definition}

\begin{remark}
To clearly visualize Proposition~\ref{prop:accordionsSlaloms}, we follow the convention to draw the $\dissection$-accordions (resp.~$\dissection\dual$-slaloms) so that they intersect the arcs of~$\dissection$ (resp.~of~$\dissection\dual$) only at the intersection points of the arcs of~$\dissection$ with the arcs of~$\dissection\dual$.
\vincent{That is interior points}
\end{remark}

\begin{remark}
The $\dissection$-slaloms and the $\dissection\dual$-accordions are defined dually and also coincide.
\end{remark}

\begin{remark}
A quick disclaimer about accordion complexes and slalom complexes of arbitrary dissections.
As stated in Proposition~\ref{prop:conditionsDualDissections}, a dissection~$\dissection$ with a face containing more than one boundary edge does not admit a dual cellular dissection~$\dissection\dual$.
This prevents to use the definitions given in this section.
Although its accordion complex (resp.~slalom complex) could still be defined as in~\cite{MannevillePilaud-accordion} (resp.~\cite{GarverMcConville}) for the disk, the resulting complexes would be join of smaller accordion complexes (resp.~slalom complexes) defined in this paper.
See \cite[Prop.~2.4]{MannevillePilaud-accordion} for a detailed statement in the case of the disk.
\vincent{A retravailler}
\end{remark}

\section{Non-kissing versus non-crossing}

In this section, we show that non-kissing complexes and non-crossing complexes actually coincide.
Namely, any pair of dual cellular dissections~$(\dissection, \dissection\dual)$ defines a locally gentle bound quiver~$\bar Q_{\dissection, \dissection\dual}$ such that the non-crossing complex~$\NCC$ is isomorphic to the non-kissing complex~$\NKC[\bar Q_{\dissection, \dissection\dual}]$.
Conversely, any locally gentle bound quiver~$\bar Q$ gives rise to a surface~$\surface_{\bar Q}$ equipped with a pair of dual cellular dissections~$\dissection_{\bar Q}$ and~${\dissection_{\bar Q}}\dual$ such that the non-kissing complex~$\NKC$ is isomorphic to the non-crossing complex~$\NCC[\dissection_{\bar Q}, {\dissection_{\bar Q}}\dual]$.
In fact, we show that the accordion, slalom, and non-kissing complexes can all be pictured in a unique way on the surface.
Interestingly, via the correspondence between pairs of dual cellular dissections and locally gentle bound quivers, the duality between the dissections~$\dissection$ and~$\dissection\dual$ translates to the Koszul duality between the locally gentle algebras of~$\bar Q$ and~$\bar Q\dual$.

\vincent{This section is just sketched at the moment.}

\subsection{The bound quiver of a dissection}

Let~$\dissection$ and~$\dissection\dual$ be two dual cellular dissections of a marked surface~$(\surface, V\sqcup V\dual)$.
Let $B$ be the set of blossom points.

\begin{definition}
\label{def:quiverDualDissections}
The bound quiver of the dissection~$\dissection$ is the bound quiver $\bar Q_{\dissection} = (Q_{\dissection}, I_{\dissection})$ defined as follows:
\begin{itemize}
 \item the set of vertices of~$Q_{\dissection}$ is the set or edges of $\dissection$;
 \item there is an arrow from~$a$ to~$b$ for each common endpoint~$v$ of~$a$ and~$b$ such that~$b$ comes immediately after~$a$ in the clockwise order around~$v$;
 \item the ideal~$I_{\dissection}$ is generated by the paths of length two in~$Q_{\dissection}$ obtained by composing arrows which correspond to triples of consecutive edges in a face of~$\dissection$.
\end{itemize}
The bound quiver of the dissection~$\dissection\dual$ is the bound quiver $\bar Q_{\dissection\dual} = (Q_{\dissection\dual}, I_{\dissection\dual})$ defined by replacing~$\dissection$ by~$\dissection\dual$ in the above.
\end{definition}

\begin{lemma}
\label{lemm:quiverOfDissectionIsLocallyGentle}
The bound quiver~$\bar Q_{\dissection} = (Q_{\dissection}, I_{\dissection})$ is a locally gentle bound quiver
\end{lemma}
\begin{proof}
Each arc~$a$ has two endpoints, and at each endpoint there is at most one arc coming immediately after~$a$ in the clockwise order, and at most one coming immediately before.
Thus, in~$Q_{\dissection}$, there are at most two arrows leaving~$a$ and at most two arriving in~$a$.
Among these arrows, those arising from the same endpoint of~$a$ will be composable, and those arising from different endpoints will yield relations.
Thus $\bar Q_{\dissection}$ is locally gentle.
\end{proof}

\begin{figure}[t]
	\capstart
	\centerline{\includegraphics[scale=.7]{quiversDissections1}}
	\medskip
	\centerline{\includegraphics[scale=.7]{quiversDissections2}}
	\caption{The quivers associated to the dissections of \fref{fig:dissections}. As stated in Proposition~\ref{prop:dualityKoszul1}, dual dissections give rise to koszul dual quivers.}
	\label{fig:quiversDissections}
\end{figure}

%Moreover, the vertices corresponding to boundary edges of~$\dissection$ are blossoming vertices.
%\pierreguy{Be more precise on the ``are blossiming vertices'' part of this sentence.}
\vincent{Explain also that we can get the blossoming quiver using the blossom points.}

\begin{definition}
\label{defi:koszulDual}
The \defn{Koszul dual} of a locally gentle bound quiver~$\bar Q = (Q,I)$ is the bound quiver~$\bar Q\koszul = (Q\koszul, I\koszul)$ defined as follows:
\begin{itemize}
 \item the quiver~$Q\koszul$ is equal to the opposite quiver of~$Q$, that is, the quiver obtained from~$Q$ by reversing all arrows;
 \item the ideal~$I\koszul$ is generated by the opposites of the paths of length two in~$Q$ that do not appear in~$I$.
\end{itemize}
\end{definition}
It is seen in \cite{BessenrodtHolm} that the (ungraded version of the) Koszul dual of a locally gentle algebra~$kQ/I$ is isomorphic to~$kQ\koszul/I\koszul$.
Note that~$\bar Q\koszul$ is gentle, and that Koszul duality is an involution and commutes with blossoming:~$(\bar Q\koszul)\koszul = \bar Q$ and~$(\bar Q\koszul)\blossom = (\bar Q\blossom)\koszul$.
Examples are represented on \fref{fig:koszulQuivers}.

\begin{figure}[t]
	\capstart
	\centerline{\includegraphics[scale=.6]{koszulQuivers}}
	\caption{The Koszul duals of the (blossoming) quivers of \fref{fig:quivers}.}
	\label{fig:koszulQuivers}
\end{figure}

\begin{proposition}
\label{prop:dualityKoszul1}
Let~$\dissection$ and~$\dissection\dual$ be two dual cellular dissections of a marked surface~$(\surface, V\sqcup V\dual)$.
The bound quivers~$\bar Q_{\dissection}$ and~$\bar Q_{\dissection\dual}$ are Koszul dual to each other.
\end{proposition}

\begin{proof}
 The duality gives a bijection~$v\in (Q_\dissection)_0\mapsto v\dual\in (Q_{\dissection\dual})_0$, and induces a bijection~$\alpha\in (Q_\dissection)_1\mapsto \alpha\dual\in (Q_{\dissection\dual})_1$ with~$s(\alpha\dual)=s(\alpha)\dual$ and~$t(\alpha\dual)=t(\alpha)\dual$.
 Moreover, for any two composable arrows~$\alpha,\beta\in(Q_\dissection)_0$, we have~$\alpha\beta\in I_\dissection$ if and only if~$s(\alpha),t(\alpha),t(\beta)$ are three consecutive edges of a~$\dissection$-cell~$f$, if and only if~$s(\alpha)\dual,t(\alpha)\dual,t(\beta)\dual$ are three consecutive arcs in~$\dissection\dual$ ending in~$f\dual$, if and only if~$\beta\dual\alpha\dual\notin I_{\dissection\dual}$.
\end{proof}

\subsection{The surface of a locally gentle bound quiver}

We now associate a surface to a locally gentle bound quiver.

\begin{definition}
\label{def:surfaceQuiver}
The surface~$\surface(\bar Q)$ of a locally gentle bound quiver~$\bar Q = (Q,I)$ is the surface obtained from the blossoming quiver~$\bar Q\blossom$ as follows:
\begin{enumerate}[(i)]
\item for each arrow~$\alpha \in Q_1\blossom$, consider a lozange~$L(\alpha)$ with four (oriented) sides
\begin{alignat*}{2}
& \Enrs{\alpha} = [v(\alpha), s(\alpha)] \qquad & \Enrt{\alpha} = [v(\alpha), t(\alpha)] \qquad & \text{(colored green)} \\
\text{and} \qquad & \Ers{\alpha} = [v\dual(\alpha), s(\alpha)] \qquad & \Ert{\alpha} = [v\dual(\alpha), t(\alpha)] \qquad & \text{(colored red)}
\end{alignat*}
placed as follows:

\centerline{
\begin{tikzpicture}[
	scale=1.5,
	thick,
	decoration={markings, mark=at position 0.5 with {\arrow{>}}},
	roundNode/.style={circle, fill=black, inner sep=2pt, outer sep=1.5pt}
	]
	\node[roundNode, label={[label distance=-3pt]left:{$s(\alpha)$}}] (s) at (-1,0) {};
	\node[roundNode, label={[label distance=-3pt]right:{$t(\alpha)$}}] (t) at (1,0) {};
	\node[roundNode, fill=green, label={[label distance=-3pt]above:{$\green v(\alpha)$}}] (v) at (0,1) {};
	\node[roundNode, fill=red, label={[label distance=-3pt]below:{$\red v\dual(\alpha)$}}] (vd) at (0,-1) {};
	\draw[color=black, postaction={decorate}] (s) -> (t) node[midway, above] {$\alpha$};
	\pic [draw, angle radius=.7cm] {angle = s--t--vd};
	\pic [draw, angle radius=.7cm] {angle = vd--s--t};
	\draw[color=green, postaction={decorate}] (v) -> (s) node[midway, above left, outer sep=-3pt] {$\green \Enrs{\alpha}$};
	\draw[color=green, postaction={decorate}] (v) -> (t) node[midway, above right, outer sep=-3pt] {$\green \Enrt{\alpha}$};
	\draw[color=red, postaction={decorate}] (vd) -> (s) node[midway, below left, outer sep=-3pt] {$\red \Ers{\alpha}$};
	\draw[color=red, postaction={decorate}] (vd) -> (t) node[midway, below right, outer sep=-3pt] {$\red \Ert{\alpha}$};
\end{tikzpicture}
}

\item for any~$\alpha, \beta \in Q_1\blossom$ with~$t(\alpha) = s(\beta)$, proceed to the following identifications:
    \begin{itemize}
    \item if~$\alpha\beta \in I$, then glue~$\Ert{\alpha}$ with~$\Ers{\beta}$,
    \item if~$\alpha\beta \notin I$, then glue~$\Enrt{\alpha}$ with~$\Enrs{\beta}$.
    \end{itemize}
\end{enumerate}
The orientations on the edges are only use to perform the identifications and can be immediately forgotten.
\end{definition}

This construction is illustrated on \fref{fig:surfaces}.
%
\begin{figure}[t]
	\capstart
	\centerline{\includegraphics[scale=.7]{surfaces}}
	\caption{The surface~$\surface_{\bar Q}$ for the quivers~$\bar Q$ of \fref{fig:quivers}. The boundaries of the surfaces are shaded, and the glue symbols indicate how to identify some edges to get the desired surfaces. The first two surfaces are disks, the third is a sphere, the fourth is a cylinder, and the fifth is a punctured torus.}
	\label{fig:surfaces}
\end{figure}

Definition~\ref{def:surfaceQuiver} constructs an orientable surface~$\surface(\bar Q)$ with boundaries.

\begin{remark}
We could equivalently construct the surface~$\surface_{\bar Q}$ as follows:
\begin{itemize}
\item consider one $(\ell+1)$-gon~$P_\omega$ for each straight walk~$\omega$ of length~$\ell$ in~$\bar Q$ or in~$\bar Q\koszul$ (just closing~$\omega$ with an additional edge connecting~$t(\omega)$ to~$s(\omega)$),
\item for each arrow~$\alpha \in Q_1\blossom$, consider the only straight walks~$\omega$ of~$\bar Q$ and~$\omega\koszul$ of~$\bar Q\koszul$ containing~$\alpha$, and glue the polygons~$P_\omega$ and~$P_{\omega\koszul}$ along their edges corresponding to~$\alpha$.
\end{itemize}
\end{remark}

The advantage of Definition~\ref{def:surfaceQuiver} is that it automatically endows with two disjoint sets~$V_{\bar Q}$ and~${V_{\bar Q}}\dual$ of marked points and two dual dissections~$\dissection_{\bar Q}$ and~${\dissection_{\bar Q}}\dual$ defined as follows.

\begin{definition}
\label{def:dissectionQuiver}
The surface~$\surface(\bar Q)$ is endowed with
\begin{itemize}
\item the set~$V_{\bar Q}$ of points~$v(\alpha)$ for~$\alpha \in Q_1\blossom$ after the identifications given by~(ii),
\item the $V_{\bar Q}$-dissection~$\dissection_{\bar Q}$ given by all sides~$\Enrs{\alpha}$ and~$\Enrt{\alpha}$ for~$\alpha \in Q_1\blossom$ after the identifications given by~(ii).
\end{itemize}
The set~${V_{\bar Q}}\dual$ and the ${V_{\bar Q}}\dual$-dissection~${\dissection_{\bar Q}}\dual$ are defined dually.
\end{definition}

\begin{proposition}
\label{prop:dissectionsAreCellular}
Let~$\bar Q$ be a locally gentle bound quiver.
Then the dissections~$\dissection_{\bar Q}$ and~$\dissection_{\bar Q}\dual$ are cellular and dual to each other.
\end{proposition}
\begin{proof}
In the construction of~$\surface(\bar Q)$, if two arrows~$\alpha$ and~$\beta$ of~$\bar Q \blossom$ are such that~$t(\alpha) = s(\beta)$ and~$\alpha\beta\in I\blossom$, then the vertices~$v(\alpha)$ and~$v(\beta)$ are identified on the surface.
Thus, any two arrows of a given path contribute the same vertex of~$V_{\bar Q}$, so~$V_{\bar Q}$ is in bijection with the set of maximal paths in~$\bar Q\blossom$.

Let~$v\in V_{\bar Q}$. The situation around~$v$ is as follows, depending on whether~$v$ corresponds to a finite or infinite maximal path of~$\bar Q\blossom$,:

\begin{figure}[h]
	\capstart
	\centerline{\includegraphics[scale=.7]{halfStar} \qquad \includegraphics[scale=.7]{star}}
	\caption{The situation around~$v$ (green vertex). On the left,~$v$ corresponds to a finite maximal path of~$\bar Q$, while on the right,~$v$ corresponds to an infinite one.}
	\label{fig:stars}
\end{figure}

In both cases, we see that~$v$ is enclosed in a polygon of~$\dissection_{\bar Q}\dual$.  This shows that~$\dissection_{\bar Q}\dual$ is a cellular dissection of~$\surface(\bar Q)$.
Dually, one shows that~$\dissection_{\bar Q}$ is also a cellular dissection.

Moreover, by construction, each edge of~$\dissection_{\bar Q}$ which is not on the boundary of~$\surface(\bar Q)$ crosses exactly one edge of~$\dissection_{\bar Q}\dual$, and vice versa.
This shows that~$\dissection_{\bar Q}$ and~$\dissection_{\bar Q}\dual$ are dual to each other.
\end{proof}

\begin{theorem}
\label{thm:bijectionLocallyGentleAndSurfaces}
 Up to isomorphism, the constructions of Definitions \ref{def:quiverDualDissections} and \ref{def:surfaceQuiver} are inverse to each other.
 They induce a bijection between the set of isomorphism classes of locally gentle algebras and the set of homeomorphism classes of marked surfaces with a pair of dual cellular dissections.
\end{theorem}
\pierreguy{How much detail for a proof? Or is it clear?}

\begin{remark}
The following observations are useful for the computation of examples.
\begin{itemize}
\item The set~$V_{\bar Q}$ has one vertex for each straight walk in~$\bar Q$ (equivalently, for each maximal path in~$\bar Q$).
      Finite straight walks yield vertices on the boundary, while infinite cyclic straight walks yield punctures.
\item The dissection~$\dissection_{\bar Q}$ has one arc for each vertex~$p \in \bar Q_0$, obtained by concataining the sides~$\Enrt{\alpha} = \Enrs{\beta}$ and~$\Enrt{\alpha'} = \Enrs{\beta'}$ where~$p = t(\alpha) = s(\beta) = t(\alpha') = s(\beta')$, $\alpha\beta \notin I$ and~$\alpha'\beta' \notin I$.
\item The dissection~$\dissection_{\bar Q}$ has one $\ell$-cell for each straight walk of length~$\ell$ in~$\bar Q\koszul$.
\item The number of boundary components of~$\surface(\bar Q)$ can be computed as follows.
      There are two natural perfect matchings whose vertices are the blossoming vertices of~$\bar Q$: one is obtained by putting an edge between two vertices~$v_1$ and~$v_2$ whenever there is a straight walk of~$\bar Q$ relating them, and the other is obtained by doing the same for~$\bar Q \koszul$.
      Let~$G$ be the superposition of these two perfect matchings.
      Then the number of boundary components of~$\surface(\bar Q)$ is the number of connected components of~$G$.
\item The genus of the surface~$\surface(\bar Q)$ is
      \[
       g = \frac{|Q_1| - |Q_0| - b + 2}{2},
      \]
      where~$b$ is the number of boundary components (see above for a way to compute~$b$).
      Indeed, ~$g=\frac{2-\chi(\surface(\bar Q))}{2}$, where~$\chi(\surface(\bar Q))$ is the Euler characteristic of~$\surface(\bar Q)$.
      The dissection~$\dissection$ provides a cellular decomposition of~$\surface(\bar Q)$ on which we fill the boundary components with disks.  
      The number of faces of this cellular decomposition is~$b$, its number of edges is~$|Q_0|$ and its number of vertices is the number of finite straight walks in~$\bar Q$, which is equal to~$2|Q_0| - |Q_1|$.
      The above formula for~$g$ follows.
\end{itemize}
A similar remark holds dually for~${V_{\bar Q}}\dual$ and~${\dissection_{\bar Q}}\dual$.
\end{remark}

\begin{proposition}
\label{prop:dualityKoszul2}
For any gentle bound quiver~$\bar Q$ with Koszul dual~$\bar Q\koszul$, the surfaces~$\surface_{\bar Q}$ and~$\surface_{\bar Q\koszul}$ coincide, but~$\dissection_{\bar Q\koszul} = {\dissection_{\bar Q}}\dual$ and~${\dissection_{\bar Q\koszul}}\dual = \dissection_{\bar Q}$.
\end{proposition}

\begin{example}[Cambrian paths]
\label{exm:CambrianPath}
\end{example}

\begin{example}[Reversed paths]
\label{exm:doublePath}
Consider the following family of locally gentle quivers:
\vincent{TODO double path}
\end{example}

\begin{example}[Double paths]
\label{exm:doublePath}
Consider the family ``double paths'' quivers indicated in \fref{fig:doublePathsQuivers}.

\begin{figure}[h]
	\capstart
	\centerline{\includegraphics[scale=.4]{doublePathsQuivers}}
	\caption{The double paths quivers with~$1, 2, 3, 4$ vertices.}
	\label{fig:doublePathsQuivers}
\end{figure}

\begin{figure}[h]
	\capstart
	\centerline{\includegraphics[scale=.7]{doublePathsSurfaces}}
	\caption{The surface~$\surface_{\bar Q}$ for the double paths with~$1, 2, 3, 4$ vertices.}
	\label{fig:doublePathsSurfaces}
\end{figure}

\end{example}

\begin{example}[Double cycles]
\label{exm:doubleCycle}
Consider the following family of locally gentle quivers:
\vincent{TODO double cycle}
\end{example}

\subsection{Non-crossing and non-kissing complexes coincide}

Let~$\bar Q$ be a locally gentle bound quiver.
For each arc of the dissection~$\dissection_{\bar Q}$ on~$\surface(\bar Q)$, we fix a point on the interior of this arc, which we call its ``middle point''.
To each walk on~$\bar Q$, we will now associate a~$\dissection_{\bar Q}$-accordion.

\begin{definition}
\label{def:curveOfAnArrow}
For any arrow~$\alpha$ on $\bar Q\blossom$, let~$\curveof(\alpha)$ be the curve on~$\surface(\bar Q)$ which goes from the middle point of the arc corresponding to $s(\alpha)$ to the middle point of the arc corresponding to $t(\alpha)$ by following the angle corresponding to~$\alpha$.
Define~$\curveof(\alpha^{-1})$ to be~$\curveof(\alpha)^{-1}$.
\end{definition}

\begin{definition}
\label{def:curveOfAWalk}
Let~$\omega = \ldots \alpha_{i-1}^{\varepsilon_{i-1}}\alpha_{i}^{\varepsilon_i}\alpha_{i+1}^{\varepsilon_{i+1}} \ldots$ be a (possibly infinite) walk on~$\bar Q$. 
Define the curve~$\curveof(\omega)$ to be the concatenation of the curves~$\curveof(\alpha_i{\varepsilon_{i}})$ of Definition \ref{def:curveOfAnArrow}.
\end{definition}

In practice, we will represent~$\curveof(\omega)$ by a curve which intersects itself only transversaly, and such that if it circles infinitely around a puncture, then it spirals towards it.

\begin{lemma}
\label{lemm:curveOfAWalkIsAccordion}
Let~$\omega$ be a walk on~$\bar Q$.  Then~$\curveof(\omega)$ is a~$\dissection_{\bar Q}$-accordion.
\end{lemma}
\begin{proof}
 This is because~$\curveof(\omega)$ follows angles in~$\dissection_{\bar Q}$ by definition.
\end{proof}

\begin{lemma}
\label{lemm:accordionsAreCurvesOfWalks}
Let~$\gamma$ be a~$\dissection_{\bar Q}$-accordion.  There exists a unique undirected walk~$\walk(\gamma)$ such that we have~$\curveof(\walk(\gamma)) = \gamma$.
\end{lemma}
\begin{proof}
 \pierreguy{TO DO}
\end{proof}

The above implies the following.

\begin{proposition}
\label{prop:walks=arcs}
The maps~$\curveof(-)$ and~$\walk(-)$ induce mutually inverse bijections between the set of undirected walks on~$\bar Q$ and the set of~$\dissection_{\bar Q}$-accordions on~$\surface(\bar Q)$.
\end{proposition}

\begin{lemma}
\label{lem:nonKissing=nonCrossing}
Non-kissing walks = non-crossing $\dissection$-accordion curves.
\end{lemma}

\begin{theorem}
\label{thm:complexesCoincide}
\begin{itemize}
\item For any locally gentle bound quiver~$\bar Q$, the non-kissing complex~$\NKC$ is isomorphic to the non-crossing complex~$\NCC[\dissection_{\bar Q}, {\dissection_{\bar Q}}\dual]$, therefore to the accordion complex~$\AC[\dissection_{\bar Q}]$ and to the slalom complex~$\SC[{\dissection_{\bar Q}}\dual]$.
\item For any pair of dual cellular dissections~$\dissection, \dissection\dual$ of an oriented surface, the non-crossing complex~$\NCC$ is isomorphic to the non-kissing complex~$\NKC[\bar Q_{\dissection, \dissection\dual}]$.
\end{itemize}
\end{theorem}

\section{Properties of the non-crossing and non-kissing complexes}

\subsection{Some technical properties of non-kissing walks}

\begin{lemma}\label{lem:nkCycles}
 Let~$\omega$ be a non-kissing walk that contains an oriented cycle~$c$.
 Then~$\omega$ is of the form~$\sigma c^\infty$ or~$c^\infty\sigma$, for some substring~$\sigma$.
\end{lemma}

\begin{proof}
 Let~$c=\alpha_1\cdots\alpha_r$ be an oriented cycle in some walk~$\omega$.
 Assume that~$\omega$ is not of form~$\sigma c^\infty$ or~$c^\infty\sigma$.
 Then, up to cyclic reordering of the arrows in~$c$, it contains a substring of the form~$\beta^\pm c^l \alpha_1\cdots\alpha_j\gamma^\pm$, for some arrows~$\beta,\gamma$ not in~$c$, some~$l\geq 1$, and some~$j<r$.
 Since~$\bar Q$ is locally gentle, this substring is~$\beta^{-1} c^l \alpha_1\cdots\alpha_j\gamma^{-1}$.
 Hence, the walk~$\omega$ contains the two substrings~$\beta^{-1}\alpha_1\cdots\alpha_j\alpha_{j+1}$ and~$\alpha_r\alpha_1\cdots\alpha_j\gamma^{-1}$, and thus self-kisses.
\end{proof}

\subsection{The countercurrent order}
The countercurrent order was introduced in~\cite{McConville} for grids and defined in \cite{PaluPilaudPlamondon} for gentle algebras.
We now adapt it to locally gentle algebras.

\begin{definition}
A \defn{marked walk}~$\omega_\star$ is a walk~$\omega$ together with a marked occurence of an arrow~$\alpha^{\pm}$ in~$\omega$.
In the case where~$\omega$ is an infinite straight walk (so~$\omega$ is a cycle which repeats infinitely on both sides), we consider that all marked occurences of an arrow~$\alpha^\pm$ are equal.
Otherwise, if~$\omega$ contains several occurrences of~$\alpha^\pm$, only one occurrence is marked.
\end{definition}

\begin{definition}
For any arrow~$\alpha \in Q\blossom_1$ and any two distinct non-kissing walks~$\mu_\star, \nu_\star$ marked at an occurrence of~$\alpha^\pm$,
let~$\sigma$ denote their maximal common substring containing that occurrence of~$\alpha$.
Since~$\mu_\star \ne \nu_\star$, this common substring~$\sigma$ is strict, so~$\mu_\star$ and~$\nu_\star$ split at one endpoint of~$\sigma$.
The \defn{countercurrent order at~$\alpha$} is defined by~$\mu_\star \prec_\alpha \nu_\star$ when~$\mu_\star$ enters and/or exits~$\sigma$ in the direction pointed by~$\alpha$, while~$\nu_\star$ enters and/or exits~$\sigma$ in the direction opposite to~$\alpha$.
\end{definition}
\pierreguy{Add a picture here?}

\begin{remark}
\label{rema:countercurrentOrderOnSurface}
Using Theorem \ref{thm:complexesCoincide}, we can translate the definition of the countercurrent order~$\prec_\alpha$ to the language of~$\dissection$-accordions.
An arrow~$\alpha$ in~$\bar Q\blossom$ corresponds to an angle at a point $v$ between two arcs of~$\dissection$.
Two distinct non-kissing walks~$\mu_\star, \nu_\star$ marked at~$\alpha$ in~$\bar Q$ correspond to two distinct non-crossing~$\dissection$-accordions~$\gamma$ and~$\delta$ passing in front of this angle.
Then~$\gamma \prec_\alpha \delta$ if~$\gamma$ passes closer to~$v$ than~$\delta$ does.
\pierreguy{Should we prove this?}
%
\begin{figure}[h]
	\capstart
	\centerline{\includegraphics[scale=.7]{orderCurves}}
	\caption{The order~$\prec_\alpha$ is given by the order in which the curves cross the angle corresponding to~$\alpha$.}
	\label{fig:orderCurves}
\end{figure}
\end{remark}

For any face~$F$ of~$\NKC$ and any arrow~$\alpha\in Q_1\blossom$, we denote by~$F_\alpha$ the set of walks of~$F$ marked at an occurence of~$\alpha^\pm$.

\begin{lemma}
\label{lemm:countercurrentOrderIsTotal}
For any face~$F$ of~$\NKC$ and any arrow~$\alpha\in Q_1\blossom$, the countercurrent order~$\prec_\alpha$ defines a total order on~$F_{\alpha}$.
\end{lemma}
\begin{proof}
 The proof of \cite[Lemma 2.21]{PaluPilaudPlamondon} applies here.
\end{proof}

\begin{lemma}
\label{lemm:facesHaveMaximalElement}
For any finite face~$F$ of~$\NKC$ and any arrow~$\alpha\in Q_1\blossom$, the set~$F_\alpha$ is either empty or admits a maximal element for the countercurrent order~$\prec_\alpha$.
\end{lemma}
\begin{proof}
 We provide two proofs of the lemma.  One is combinatorial and applies to~$\NKC$, and the other is geometric and applies to~$\NCC[\dissection_{\bar Q}, {\dissection_{\bar Q}}\dual]$, using the isomorphism of Theorem \ref{thm:complexesCoincide} and Remark \ref{rema:countercurrentOrderOnSurface}.

 {\it Combinatorial proof.}
 Note that even though~$F$ is finite, the set of marked walks~$F_\alpha$ might be infinite if~$\alpha$ belongs to an oriented cycle of~$Q$.
 Assume that~$F_\alpha$ is non-empty, and infinite.
 Any walk that contains an infinite number of copies of~$\alpha$ can be written in the form~$\omega=\sigma c^\infty$ where~$\sigma$ is a (possibly infinite) substring and~$c$ is an oriented cycle containing~$\alpha$.
 Then~$\omega$ marked at the first occurrence of~$\alpha$ in~$c^\infty$ is greater than~$\omega$ marked at any other occurrence of~$\alpha$ in~$c^\infty$.
 We can thus safely remove all copies of~$\omega$ marked at an occurrence of~$\alpha$ in~$c^\infty$ but the first one.
 After applying this procedure a finite number of times,~$F_\alpha$ becomes finite, and the claim follows.

 {\it Geometric proof.} View~$F$ as a face of the non-crossing complex~$\NCC[\dissection_{\bar Q}, {\dissection_{\bar Q}}\dual]$.
 The arrow~$\alpha$ then corresponds to an angle in~$\dissection$ at a vertex~$v$.

 For any curve~$\gamma$ in~$F$, even though~$\gamma$ may pass through the above angle infinitely many times, there is one instance that is furthest from the angle:
 indeed, if ~$\gamma$ passes infinitely many times through~$\alpha$, then~$v$ is a puncture and~$\gamma$ circles infinitely towards it.
 But then~$\gamma$ passes through~$\alpha$ only finitely many times before it starts circling~$v$ (in either direction), and one of these times is the furthest from~$v$.

 Therefore, for every curve~$\gamma\in F$, there is a maximal occurence of~$\alpha$ in~$\gamma$.
 Since~$F$ is finite, the set of these maximal occurences is finite, so it admits a maximal element.  
 This element is the maximal element of~$F_\alpha$.
\end{proof}

\begin{lemma}\label{lem:ExistenceOfDistinguishedArrows}
Let~$F$ be a finite face of~$\NKC$, let~$\omega$ be a walk in~$F$, and let~$\alpha$ be an arrow on~$\omega$.
Assume that~$\omega$ is not an infinite straight walk.
Then, there is an arrow~$\beta$ oriented in the same direction as~$\alpha$ in~$\omega$ such that~$\omega$ (marked at~$\beta$) is maximal in~$F$ for~$\prec_\beta$.
\end{lemma}

\begin{proof}
Let~$\omega$ be a walk in some non-kissing facet~$F$, marked at an arrow~$\alpha$.
If~$\omega$ is not maximal at~$\alpha$, there is a walk~$\mu_0\in F$ such that~$\omega\prec_\alpha \mu_0$.
Let~$\sigma_0$ be the maximal common substring of~$\omega$ and~$\mu_0$ containing~$\alpha$.
Since~$\omega$ and $\mu_0$ are different, they split at some endpoint of~$\sigma_0$.
At this endpoint, the arrow~$\beta$ of~$\omega$ not in~$\sigma_0$ points in the same direction as~$\alpha$ since~$\omega \prec_\alpha \mu_0$.
If~$\omega$ is maximal at~$\beta$ in~$F$, we are done.
Otherwise, let~$\mu_1$ be maximal in~$F$ for~$\prec_\beta$.
Since~$\mu_1$ does not kiss with~$\mu_0$, and~$\omega\prec_\beta\mu_1$,~$\mu_1$ contains~$\sigma_0$.
Let~$\sigma_1$ be the maximal common substring of~$\omega$ and~$\mu_1$ containing~$\beta$.
As remarked above,~$\sigma_1$ contains~$\sigma_0$.
Since~$\omega$ is not maximal at~$\beta$, the walks~$\omega$ and~$\mu_1$ split at some endpoint of~$\sigma_1$.
By Lemma\ref{lem:nkCycles}, Iterating this procedure eventually stops ($\mu_j=\omega$ for some $j$) unless~$\omega$ is an infinite straight walk.
\end{proof}

\begin{lemma}
\label{lemm:facesAreBounded}
There is a bound on the number of elements of a finite face of~$\NKC$.
\end{lemma}
\begin{proof}
 We provide two proofs of the lemma.
 One is combinatorial and applies to~$\NKC$, and the other is geometric and applies to~$\NCC[\dissection_{\bar Q}, {\dissection_{\bar Q}}\dual]$.

 {\it Combinatorial proof.} By Lemma~\ref{lem:ExistenceOfDistinguishedArrows}, the number of walks in a finite face of~$\NKC$ is bounded by~$|Q_1\blossom|+s_\infty$, where~$s_\infty$ is the number of infinite straight walks (i.e. the number of oriented cycles in~$Q$).

 {\it Geometric proof.} Let~$F$ be a face of the non-crossing complex~$\NCC[\dissection_{\bar Q}, {\dissection_{\bar Q}}\dual]$.
 Its elements are pairwise compatible~$\dissection$-accordions on the surface~$\surface(\bar Q)$.
 Thus~$F$ can be viewed as a partial tagged triangulation of~$\surface(\bar Q)$, by letting the~$\dissection$-accordions circling a puncture in (say) clockwise order be seen as ``notched'' arcs to said puncture, and the ones circling it in counter-clockwise orientation be ``un-notched'' arcs.
 It is known (see, for instance, \cite{FominShapiroThurston}) that any such partial tagged triangulation can be embedded into a full triangulation, and that triangulations contain a finite number of arcs.
 Thus~$F$ is finite, and is contained in a finite maximal face.
\end{proof}

\begin{lemma}
\label{lemm:facetsAreFinite}
Any face of~$\NKC$ and~$\NCC[\dissection_{\bar Q}, {\dissection_{\bar Q}}\dual]$ is contained in a finite maximal face.
\end{lemma}
\begin{proof}
 If~$\NKC$ admits an infinite face, then all subsets of this face are also faces, so~$\NKC$ admits faces of arbitrary finite cardinality.
 This contradicts Lemma \ref{lemm:facesAreBounded}.
\end{proof}

\subsection{Purity of the non-crossing and non-kissing complexes}

A simplicial complex is \defn{pure of dimension~$d$} if all its maximal faces have dimension~$d$.
The aim of this section is to prove the following purity result.

\begin{proposition}
\label{prop:purity}
 The the non-kissing complex~$\NKC$ and the non-crossing complex~$\NCC[\dissection_{\bar Q}, {\dissection_{\bar Q}}\dual]$ are pure of dimension~$3|Q_0| - |Q_1|$.
\end{proposition}

Note that Proposition \ref{prop:purity} was proved for~$\NKC$ in \cite[Corollary 2.27]{PaluPilaudPlamondon} in the case where~$kQ/I$ is finite-dimensional.
Note also that, by Theorem \ref{thm:complexesCoincide} above,~$\NKC$ and~$\NCC[\dissection_{\bar Q}, {\dissection_{\bar Q}}\dual]$ coincide, so it suffices to prove the Proposition for one of them.

The following definition from \cite[Definition 2.23]{PaluPilaudPlamondon} extends to the setting locally gentle algebras.

\begin{definition}
\label{defi:distinguishedWalksAndArrows}
Let~$\alpha \in Q\blossom_1$ be an arrow and let~$F$ be a face of~$\NKC$ containing the straight walk passing through~$\alpha$.
Consider the marked walk~$\omega_\star \eqdef \max_{\prec_\alpha} F_\alpha$ (which exists by Lemma \ref{lemm:facesHaveMaximalElement}),
and let~$\omega$ be the walk obtained by forgetting the marked arrow in~$\omega_\star$. 
We say that~$\omega$ is the \defn{distinguished walk} of~$F$ at~$\alpha$ and that~$\alpha$ is a \defn{distinguished arrow} on~$\omega \in F$.
We will use the notation
\[
\distinguishedWalk{\alpha}{F} \eqdef \max\nolimits_{\prec_\alpha} F_\alpha
\qquad\text{and}\qquad
\distinguishedArrows{\omega}{F} \eqdef \set{\alpha \in \omega}{\distinguishedWalk{\alpha}{F} = \omega}.
\]
\end{definition}

The following result extends \cite[Proposition 2.26]{PaluPilaudPlamondon} to the case of locally gentle algebras.

\begin{proposition}
\label{prop:numberOfDistinguishedArrows}
Let~$F$ be a facet of~$\NKC$.
\begin{itemize}
 \item Each bending walk of~$F$ has exactly two distinguished arrows.
 \item Each finite straight walk of~$F$ has exactly one distinguished arrows.
 \item Each infinite straight walk of~$F$ has exactly zero distinguished arrows.
\end{itemize}
\end{proposition}
\begin{proof}
 \pierreguy{Vincent has the proof of this proposition.}
\end{proof}

We can now prove the purity result.

{\it Proof of Proposition \ref{prop:purity}.}
\pierreguy{TO DO}

\subsection{Thinness of the non-crossing and non-kissing complexes, mutation}

A simplicial complex is \defn{thin} if each of its codimension~$1$ faces is contained in exaclty two facets.
We state the following results, which generalize \cite[Proposition 2.31 and Corollary 2.33]{PaluPilaudPlamondon}, whose proofs apply withouth change.

\begin{proposition}
\label{prop:flip}
Let~$F$ be a non-kissing facet in~$\NKC$ and let~$\omega \in F$ be a bending walk.
Let~$\alpha$ and~$\beta$ be the distinguished arrows of~$\omega$ (see Proposition \ref{prop:numberOfDistinguishedArrows}), and~$\sigma$ be the distinguished substring of~$\omega$, which splits~$\omega$ into~$\omega = \rho \sigma \tau$.
Let~$\alpha'$ and~$\beta'$ be the other two arrows of~$Q_1\blossom$ incident to the endpoints of~$\sigma$ and such that~$\alpha'\alpha \in I$ or~$\alpha\alpha' \in I$, and $\beta'\beta \in I$ or~$\beta\beta' \in I$.
Let~$\mu \eqdef \distinguishedWalk{\alpha'}{F \ssm \{\omega\}}$ and~$\nu \eqdef \distinguishedWalk{\beta'}{F \ssm \{\omega\}}$ be the distinguished walks of~$F \ssm \{\omega\}$ at~$\alpha'$ and~$\beta'$ respectively.
Then
\begin{enumerate}[(i)]
\item The walk~$\mu$ splits into~$\mu = \rho' \sigma \tau$ and the walk~$\nu$ splits into~$\nu = \rho \sigma \tau'$.
\item The walk~$\omega' \eqdef \rho' \sigma \tau'$ is kissing~$\omega$ but no other~walk~of~$F$. Moreover, $\omega'$ is the only other (undirected) walk besides~$\omega$ which is not kissing any other walk of~$F \ssm \{\omega\}$.
\end{enumerate}
\end{proposition}

\begin{corollary}
The non-kissing complex~$\RNKC$ is pure of dimension~$|Q_0|$.  Moreover, it is thin and is a pseudomanifold without boundary.
\end{corollary}

\vincent{Orient the flip graph. Connection to the lattice of torsion classes?}

\section{Representation-theoretic interpretation}

\subsection{$\dissection$-accordions and string modules}

\subsection{$\dissection\dual$-slaloms and minimal projective presentations}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{alpha}
\bibliography{accordionComplexSurfaces}
\label{sec:biblio}

\end{document}