### 2018-07-12 firstPaper/firstPaperLong/figures/gcmatrices.tex

\(
\begin{blockarray}{ccccccc}
	& {\color{red} \bullet} & {\color{orange} \bullet} & {\color{yellow} \bullet} & {\color{green} \bullet} & {\color{blue} \bullet} & {\color{violet} \bullet} \\
	\begin{block}{c(cccccc)}
	1 & 0 & 0 & 0 & 0 & 0 & \!\!-1\\
	2 & 0 & 0 & 0 & 0 & \!\!-1\!\! & 0 \\
	3 & 0 & 1 & 0 & 1 & 0 & 0 \\
	4 & 0 & 0 & 0 & \!\!-1\!\! & 0 & 0 \\
	5 & 0 & 0 & 1 & 1 & 1 & 0 \\
	6 & 1 & 0 & 0 & 0 & 0 & 0 \\
	\end{block}
	& & & \multicolumn{2}{c}{$\gvectors{F\ex}$} & &
\end{blockarray}
%
\qquad
%
\begin{blockarray}{ccccccc}
	& {\color{red} \bullet} & {\color{orange} \bullet} & {\color{yellow} \bullet} & {\color{green} \bullet} & {\color{blue} \bullet} & {\color{violet} \bullet} \\
	\begin{block}{c(cccccc)}
	1 & 0 & 0 & 0 & 0 & 0 & \!\!-1 \\
	2 & 0 & 0 & 1 & 0 & \!\!-1\!\! & 0 \\
	3 & 0 & 1 & 0 & 0 & 0 & 0 \\
	4 & 0 & 1 & 1 & \!\!-1\!\! & 0 & 0 \\
	5 & 0 & 0 & 1 & 0 & 0 & 0 \\
	6 & 1 & 0 & 0 & 0 & 0 & 0 \\
	\end{block}
	& & & \multicolumn{2}{c}{$\cvectors{F\ex}$} & &
\end{blockarray}
\)