### 2018-07-13 firstPaper/corrections/replyToReferee.txt

We thank the referee for their helpful comments.  We have tried to implement them
all.  Below are our replies to each comment.  We have also made a few additional
changes, which we first list.

(a) In Notation 1.50, the equalities were wrong and have been corrected.

(b) We have added some details (in a new Proposition 2.24) explaining why the 
    non-kissing facets are finite. In doing so, we had to slightly reorganise 
    Subsection 2.2.2 (new Lemma 2.23 was previously part of the proof of 
    Proposition (renumbered) 2.28).

* In the introduction, and in the introductory paragraphs of several sections
  the authors are sometimes quite abrupt and discuss notions that have not been
  introduced yet. For example examples this happens in
  - page 1, line 2: non-kissing paths
  - page 43: $\mathcal{L}_{nk}(\overline{Q})$
  Could this notions be introduces, at least vaguely, before using them?

  Reply: page 43: We have added a reference to the definition.

* page 1 paragraph 1: ``The interest in'' rather then ``The interest for''

  Reply: Done.

* page 2 par 3: ``description of flips''

  Reply: Done.

* par: 3 ``remarquable'' is French

  Reply: Done.

* page 3: the footnotes refer to the table on next page; they are quite
  confusing here.

  Reply: Done.

* page 6, last line: while the meaning of $I^{rev}$ is clear this has not been
  introduced formally.

  Reply:

* Figure 1: all other figures in the paper are not separated into sub figures by
  lines as this one is

  Reply: Done.

* Def 1.12: say ``The algebra ... associated to a gentle bound quiver is called''
  or something similar.

  Reply: Done.

* page 15 above 1.1.7: all of a sudden there is a shorthand for band modules
  that was not introduced

  Reply: Corrected.

* page 15 bottom: this formula looks quite ugly, can you remove the text from
  it by defining the two sets first? Even better can you replace the formula 
  with a sentence?

  Reply: The formula has been replaced by a sentence.

* page 17: the name ``dance'' suggests a symmetric notion but it is not. If
  there in no better name available it should at least be stated clearly that
  this is not a symmetric notion.

  Reply: We have replaced "dances with" with "reaches for".

* page 24 intro: ``the operation of flip''

  Reply: Done

* Example 2.2 and many more occurrences: Through the paper there is a recurring
  example and the authors decided to name it by $Q_{exm}$ et similia.
  This makes it look like it is some sort of notation when it is not. I would
  suggest to remove this convention and to just refer to it as the recurring
  example or the example in figure 15.

  Reply: We have removed all appearances of _{exm}, and replaced them by references to the relevant figures.

* Just below: ``See Figure 15'' is not needed here

  Reply: Removed.

* bottom of page 24 what is a non-kissing cycle? Do you mean not self-kissing?

  Reply: That was a mistake. Thank you for spotting it! We meant non-oriented cycle.

* Figure 17: $Q_{pa}$ as above why is an example given a notation-like name?

  Reply: $Q_pa$ is now replaced by the more traditional $\overrightarrow{A_n}$.

* Figure 18: crossing walks have not been introduced

  Reply: We have replaced "crossing" by "non-kissing and having a common substring".  
         The term "crossing walks" appeared nowhere else in the paper.

* remark 2.20: the first bullet point ``should'' should the ``then'' you are
  stating a fact not an hope

  Reply: Fixed.

* proof of 2.21: ``follow'' is not really clear here

  Reply: Clarified.

* def 2.36: is any flip either increasing or decreasing?

  Reply: Yes. This has now been made axplicit and we have added the definition of decreasing.

* def 3.4: do you mean ``any strict subset''?

  Reply: Yes, fixed.

* def 3.16 and others: $\mathcal{S}$ and $S$ are not easily distinguishable

  Reply:

* Proof of theorem 3.20: the numberings do not correspond to each other

  Reply: The numberings have been corrected.

* page 49 previous to last paragraph: ``ordered by:'' should be ``ordered by''

  Reply:

* Example 3.36: Figure 34 does not illustrate $\omega$

  Reply:

* Example 3.44: It appears to me that the map, which is not explicit, is on the
  quotient of $Bic$

  Reply:

* pages 59,60,64,65,68,73 and probably more: the text goes outside the margin

  Reply: Fixed.

* remark 3.72: do you mean ``Figure 36 (bottom)''?

  Reply:

A note about notation. This article uses a quite creative notation set. For
example the blossoming quiver has a little flower as exponent and the operation
of blinding is represented by two half circle to suggest blinkers. I am not sure
I am happy about these choices but I do not have better suggestions to offer.

  Reply: While we agree that the notation set is non-conventional, we believe
         that it fits well with our chosen terminology.  We thus respectfully 
         propose to keep the notation as it is.