### 2020-06-29 firstPaper/exposeFPSAC1807/figures/gcmatrices.tex

\(
\begin{blockarray}{cccccccc}
	& & {\color{red} \bullet} & {\color{orange} \bullet} & {\color{yellow} \bullet} & {\color{green} \bullet} & {\color{blue} \bullet} & {\color{violet} \bullet} \\
	\begin{block}{cc(cccccc)}
	1 & & 0 & 0 & 0 & 0 & 0 & \!\!-1\\
	2 & & 0 & 0 & 0 & 0 & \!\!-1\!\! & 0 \\
	3 & & 0 & 1 & 0 & 1 & 0 & 0 \\
	4 & & 0 & 0 & 0 & \!\!-1\!\! & 0 & 0 \\
	5 & & 0 & 0 & 1 & 0 & 1 & 0 \\
	6 & & 1 & 0 & 0 & 0 & 0 & 0 \\
	\end{block}
	& & & & \multicolumn{2}{c}{$\gvectors{F}$} & &
\end{blockarray}
%
\qquad
%
\begin{blockarray}{cccccccc}
	& & {\color{red} \bullet} & {\color{orange} \bullet} & {\color{yellow} \bullet} & {\color{green} \bullet} & {\color{blue} \bullet} & {\color{violet} \bullet} \\
	\begin{block}{cc(cccccc)}
	1 & & 0 & 0 & 0 & 0 & 0 & \!\!-1 \\
	2 & & 0 & 0 & 1 & 0 & \!\!-1\!\! & 0 \\
	3 & & 0 & 1 & 0 & 0 & 0 & 0 \\
	4 & & 0 & 1 & 1 & \!\!-1\!\! & 0 & 0 \\
	5 & & 0 & 0 & 1 & 0 & 0 & 0 \\
	6 & & 1 & 0 & 0 & 0 & 0 & 0 \\
	\end{block}
	& & & & \multicolumn{2}{c}{$\cvectors{F}$} & &
\end{blockarray}
\)